\documentclass[12pt]{article}

% Shared packages/macros/bibliography setup
\input{sections/preamble}

\title{Incentives to Coordinate: Elite Endorsements in Primary Elections}
\author{Katherine Parslow}
\date{Last Updated: April 22, 2026}

\begin{document}
\maketitle

\begin{abstract}
Political elites use endorsements to shape voter beliefs and influence primary election outcomes, yet endorsement patterns vary across elections. Some elections see elites rapidly coalesce around a particular candidate, while others remain fragmented, prolonging competition. To understand this variation, I develop a formal model where elites must strategically balance ideological preferences over candidates with the political returns of backing the winning candidate. I consider two distinct structures for political rewards: fixed rewards, where benefits are constant across endorsers, and split rewards, where political benefits are divided amongst the winning candidate's supporters. Equilibrium analysis reveals that fixed rewards encourage strategic complementarity between endorsements, fostering elite coordination. Conversely, split rewards encourage elites to splinter and support different candidates. Additionally, electoral uncertainty intensifies fragmentation by raising elites' ideological thresholds for coordination. These findings help explain observed variation in endorsement patterns across primaries and suggest that differences in elite incentives critically shape the nomination process.
\end{abstract}

\input{sections/introduction}
\input{sections/literature-review}
\input{sections/model}
\input{sections/analysis}

\printbibliography
\end{document}
